---
title: "DGE Enrichment Analysis & GSEA ranked list prep"
author: "Martin Gordon"
date: "8/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Enrichment Analysis

What pathways do the genes belong to?
Cancer pathways?
miRNA targeting the genes?

Try clusterprofiler tool and prepare input file for GSEA 

```{r}
library(org.Hs.eg.db)
library(DOSE) #Disease ontology; description of gene products with disease perspectvies
library(pathview)
library(clusterProfiler)
library(AnnotationHub)
library(ensembldb)
library(tidyverse)
library(fgsea)
```

EnrichGO Overepresentation Analysis

```{r}
#enrichGO; use entrez gene ids from sig list and compare to universe 
# read in sig_id list and extract row names from res (all genes, including those w/o significance; used as universe for hypogeometric testing)

#Overrepresentation analysis; determine whether known biological functions or processes are over-represented (= enriched) in an experimentally-derived gene list, e.g. a list of differentially expressed genes (DEGs). The p-value can be calculated by hypergeometric distribution. using Fishers exact test


#Overepreesnentation Analysis (ORA) vs GSEA:
#ORA only need vector of gene IDs, GSEA needs ranked list

#sig list from DESeq2 analysis ( p.adj <= 0.05 and log2FC > =1 )

sig_genes
summary(sig_genes)

#'universe' gene set; all genes tested 

all_genes <- rownames(LFC) #pull out entrez ids 
summary(all_genes) #already character


## Run GO enrichment analysis 
ego <- enrichGO(gene = sig_genes, 
                universe = all_genes,
                keyType = "ENTREZID",
                OrgDb = org.Hs.eg.db, 
                ont = "BP", #Biological Processes GO term, also done for CC and MF
                pAdjustMethod = "BH", 
                qvalueCutoff = 0.05, 
                readable = TRUE)

cluster_BP_summary <- data.frame(ego)
View(cluster_BP_summary)

#take subset of the dataframe and plot

## Dotplot 
pdf(file="~/Documents/MSC_project/data/DESeq_output/Enrichment_Analysis/GO_enrichment_BP.pdf")
dotplot(ego, showCategory=30) + ggtitle("Dotplot for ORA")  +
  theme_classic()

dev.off()

#take subset of ego_p
ego_p <- as.data.frame(ego) 
ego_p <- ego_p[1:30,]


ggplot(ego_p, aes(x = GeneRatio, y = fct_reorder(Description, GeneRatio))) + 
               geom_point(aes(size = GeneRatio, color = p.adjust)) +
               theme_bw(base_size = 14) +
        scale_colour_gradient(limits=c(0, 0.20), low="red") +
        ylab(NULL) +
        ggtitle("GO pathway enrichment")


# enrichment GO plot; shows relationship of top x most significantly enriched GO terms (padj) by grouping similiar terms together
#colours represent p-values relative to others (brighter more significant)
pdf(file="~/Documents/MSC_project/data/DESeq_output/Enrichment_Analysis/enrichment_GO_p.pdf")
emapplot(ego, showCategory = 20)

dev.off()
```

```{r}
#Cateogry netplot; relationships between the genes associated with the top five most significant GO terms and the fold changes of the significant genes associated with these terms (color)
#size of GO terms reflects pvalues of the terms- more significant are larger
#good for hypothesis geenration; idenify genes that may be NB for some of the afffected processes

#read in the two tables of significant genes and cbind write out 

up_reg_list <- read.csv2('/Users/martingordon/Documents/MSC_project/data/DESeq_output/miR-17-92+_vs_ctrl_upreg.txt', sep = '\t', header =  T)

down_reg_list <- read.csv2('/Users/martingordon/Documents/MSC_project/data/DESeq_output/miR-17-92+_vs_ctrl_downreg.txt', sep = '\t', header =  T)

#bind together and write to new file
sig_gene_list <- rbind(up_reg_list, down_reg_list)
write.table(x = sig_gene_list, file = '~/Documents/MSC_project/data/DESeq_output/all_DEG.txt', sep ='\t', col.names = T, row.names = F)


## To color genes by log2 fold changes, we need to extract the log2 fold changes from our results table creating a named vector
OE_foldchanges <-  sig_gene_list$log2FoldChange

names(OE_foldchanges) <- sigOE$gene

## Cnetplot details the genes associated with one or more terms - by default gives the top 5 significant terms (by padj)
cnetplot(ego, 
         categorySize="pvalue", 
         showCategory = 5, 
         foldChange=OE_foldchanges, 
         vertex.label.font=6)

## If some of the high fold changes are getting drowned out due to a large range, could set a maximum fold change value

pdf(file="~/Documents/MSC_project/data/DESeq_output/Enrichment_Analysis/category_netplot.pdf")
cnetplot(ego, 
         categorySize="pvalue", 
         showCategory = 6, 
         foldChange=OE_foldchanges, 
         vertex.label.font=5,
         layout = 'kk',
         circular = F )

dev.off()
```


GSEA analysis 
Determine if a ranked gene list is significantly enriched in a prior defined biological processes/gene sets
ORA only need vector of gene IDs, GSEA needs ranked list
Create ranked list in R for use with GSEA app (ranked based on signed p-value)

```{r}
#Prepare file for GSEA input; input ALL genes for testing (including middle ranked (non-significant as removing can reduce power)
#rank genes based on signed pval

x <- LFC_df
x <- add_rownames(x, var = "entrezgene_id") #row to col 
head(x)
x$fcsign <- sign(x$log2FoldChange) #returns col of sign (+/- of LFC)
x$logP=-log10(x$pvalue) #take log of all p val
x$metric= x$logP/x$fcsign #signed pval ranking
x <- x[,c('entrezgene_id', 'metric')]
x <- x[order(-x$metric),] #order in descending order

x <-  na.omit(x) #remove rows with na 
dim(x)
x <- x[!duplicated(x$entrezgene_id),] #remove duplicates
dim(x)


#map entrez-id to symbol for use with fgsea
my.keys <- as.character(rownames(x))


ens2symbol <- AnnotationDbi::select(org.Hs.eg.db,
                                    key=my.keys, 
                                    columns="SYMBOL",
                                    keytype="ENTREZID")

ens2symbol <- as_tibble(ens2symbol)
colnames(ens2symbol) <- c("entrezgene_id", "symbol")


#join dfs
res <- inner_join(x, ens2symbol, by=c("entrezgene_id"="entrezgene_id"))

#drop rows with na
res <- na.omit(res)
dim(res) #16278 left of 16320

head(x)
 #signed pval ranking
res <- res[,c('symbol', 'metric')]
res <- res[order(-res$metric),] #order in descending order

res <-  na.omit(res) #remove rows with na 
dim(res)
res <- res[!duplicated(res$symbol),] #remove duplicates
dim(res)


#write the lists - used both as loss of some genes when mapping to gene symbols. No row or col names for .rnk file
write.table(x,file="~/Documents/MSC_project/data/DESeq_output/Enrichment_Analysis/msc_genes_entrezid.rnk",quote=F,sep="\t",row.names=F, col.names = F)

write.table(res,file="~/Documents/MSC_project/data/DESeq_output/Enrichment_Analysis/msc_genes_symbol.rnk",quote=F,sep="\t",row.names=F, col.names = F)

```


Code for plotting GSEA analysis results 
Read in results of GSEA analysis and plot significant results (p.adj < 0.05)


```{r}
pdf(file="~/Documents/MSC_project/data/DESeq_output/Enrichment_Analysis/sigKEGG_pathways.pdf")

#too many terms so just plot significant ones
keggp <- fgseaResTidy[fgseaResTidy$padj<0.05,]

ggplot(keggp, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=NES < 0)) +
  scale_fill_manual(name = '', values = setNames(c("#F46D43","#74ADD1"),c("", "")), guide = NULL) +

 # scale_fill_manual(values = c("#fa624b","#5e61fc"))
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
       title="Hallmark pathways NES from GSEA") + 
  theme_classic() 

dev.off()
```


