---
title: "Project WGCNA Analysis"
author: "Martin Gordon"
date: "8/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r}
library(WGCNA)
library(RColorBrewer)
library(gplots)
library(reshape2)
library(DESeq2)
library(ggplot2)
library(knitr)
library(RColorBrewer)
library(pvclust)
library(genefilter)
library(flashClust)
library(dplyr)
library(directlabels)

# The following setting is important, do not omit.
# Allow multi-threading within WGCNA. This helps speed up certain calculations.
# At present this call is necessary.
# Any error here may be ignored but you may want to update WGCNA if you see one.
options(stringsAsFactors = FALSE)
```

## Read in expression and metadata

```{r}
# regularised log transformed (rld) data for WGCNA (pre-trimmed to remove low counts and 0's)
#work with rld-transformed data (or logged transformed data) as WGCNA built for microarray like data (normally distributed, unlike rna-seq which is discrete overdispersed, so tranform data to approximate something more 'normal like' )

counts_data <- (assay(rld))

head(counts_data)

write.table(counts_data, file = '/Users/martingordon/Documents/MSC_project/data/WGCNA/Normalised_Counts_vst.txt', sep ='\t', col.names = T, row.names = T)


#assess sample quality

col_dendo <- c("#0072B2","#0072B2","#0072B2", "#0072B2",'#D55E00','#D55E00','#D55E00', '#D55E00')

#all samples high similarity and samples segregate by clustering, although two mutants different to others

#pdf(file='~/Documents/MSC_project/data/WGCNA/sample_clust(rld).pdf')
heatmap.2(cor(counts_data), RowSideColors=col_dendo,
          trace='none', main='Sample correlations (vst-transformed)', key.title = NA, margins = c(13,11))

#dev.off()


x = melt(as.matrix(assay(rld)))

#density plot of normalised count data

colnames(x) = c('gene_id', 'sample', 'value')
#pdf(file='~/Documents/MSC_project/data/WGCNA/sample_density(rld).pdf')
ggplot(x, aes(x=value, color=sample)) + geom_density() + ggtitle('rld transform') + theme_classic()

#dev.off()


#this removes any genes with 0 variance,contribute noise
counts_data <- counts_data[apply(counts_data, 1, var) > 0,]

#tested this, doesnt really improve model fit 

#create variable containing only top 50% most variable genes to test if better for WGCNA 
#not recommended by authors to filter by v high variance genes as may be noise
counts_data2 <- varFilter(counts_data) #filter threshold top 50% of most variable genes 
dim(counts_data2) #half of counts_data

```

```{r}
#first transpose the matrix

datExpr = as.data.frame(t(counts_data2[,]))
head(datExpr, n =8)

#row names are samples, col are the genes

#check for samples and genes with too many missing values (not expected to be many due to previous filtering)
gsg = goodSamplesGenes(datExpr, verbose = 3)
gsg$allOK #all genes and samples sufficent quality to proceede



sampleTree = hclust(dist(datExpr), method = "average");
#Plot the sample tree: 
#two clusters of 2 and 6 
sizeGrWindow(12,9)
pdf(file = '/Users/martingordon/Documents/MSC_project/data/WGCNA/rld_sample_deodnogram', width = 12, height = 9);
par(cex = 0.6);
par(mar = c(0,4,2,0))
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5,
cex.axis = 1.5, cex.main = 2)

dev.off()
```

```{r}
#read in sample data and match to the expression data
traitData = read.csv('/Users/martingordon/Documents/MSC_project/data/WGCNA/sample_table.txt', sep = '\t')

dim(traitData)
names(traitData)


#sample trait data 

Control <- c(1,1,1,1, 0,0,0,0)
miR_Overexp <- c(0,0,0,0, 1,1,1,1)

datTraits <- data.frame(Control, miR_Overexp)

# Form a data frame analogous to expression data that will hold the clinical traits.
rownames(datTraits) = sample_table$Sample;

#check pheontype data corresponds to exp data
table(rownames(datTraits) == rownames(datExpr))   
```

Network Construction; selecting a similarity metric

```{r}

#thresholding connections; all nodes connected in wgcna network; soft thresholding emphasises stronger connections relative to weaker ones 
#could also use a hard threshold if doesnt work and just select nodes and edges above the threshold

#Choosing the soft-thresholding power: analysis of network topology
#pickSoftThreshold that performs the analysis of network topology and aids the user in choosing a proper soft-thresholding power to achieve scale free topology

powers = c(c(1:10), seq(from = 12, to=20, by=2))
sft = pickSoftThreshold(datExprT, powerVector = powers, verbose = 2) 
#signed network as want to keep track of positive and negative regulation


# Plot the results:
sizeGrWindow(9, 5)
#pdf(file='~/Documents/MSC_project/data/WGCNA/soft_thresholding.pdf')
par(mfrow = c(1,2));
cex1 = 0.5;
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
labels=powers,cex=cex1,col="red");

# this line corresponds to using an R^2 cut-off of h
#selected 0.8 as threshold to 
abline(h=0.90,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")

sft

dev.off()

#looking at output for the connectivity, can see that even when raised to power of 30, R2 not .9 for model fit and mean connectivity still abouve 200
#22? above .8 and median connectivity ~ 182...
#selected power 18 for signed network R2 val of 0.80, mean k of 805.. v high connectivity with sample size and possibly heterogeneity havign a huge impact, try futher filtering?

#will try with other similarity metrics and  DEGs (not recommended by WGCNA) 
#filter the expression matrix by gene names in DEG list (wont likely be scale free, but may reduce connectivity..)

```

DEG data; testing, doesnt fit model well and connectivity remains v high
```{r}

#list of sig genes from DESeq (adj.pval 0.05, log2FC 1)
dim(norm_sig)
class(norm_sig)
norm_sig2 <- as.numeric(rownames(norm_sig)) #convert from character to numberic for filtering 


#DEG list
counts_data2 <- counts_data[rownames(counts_data) %in% norm_sig2,] #filter counts data by rowname of sig genes
dim(counts_data2)


datExprDEG = as.data.frame(t(counts_data2[,]))

powers = c(c(1:10), seq(from = 12, to=60, by=2))
sft = pickSoftThreshold(datExprDEG, powerVector = powers, networkType = "signed", dataIsExpr = T, verbose = 2) 
#signed network as want to keep track of positive and negative regulation


# Plot the results:
sizeGrWindow(9, 5)
#pdf(file='~/Documents/MSC_project/data/WGCNA/soft_thresholding(DEG).pdf')
par(mfrow = c(1,2));
cex1 = 0.5;
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
labels=powers,cex=cex1,col="red");

# this line corresponds to using an R^2 cut-off of h
#selected 0.8 as threshold to 
abline(h=0.80,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")

#reduces connectivity, but not scale free 
sft$fitIndices
```

Biweight midcorrelation; more robust to outliers than Pearson 
also called bicor; a measure of similarity between samples. It is median-based, rather than mean-based, thus is less sensitive to outliers, and can be a robust alternative
#choosing power for bicor signed network

#resons for using signed; harder to detect hubs using unsigned(what direction are they correlated with the variable?), keep positively correlated nodes seperately and gives better biological interpretation. )eigengene and gene correlation positive)


```{r}
#construct different similarity metrics for measruring scale free topology
powers = c(c(1:10), seq(from = 12, to=30, by=2))

#decrrease blocksize and try this again tomorrow
sft =pickSoftThreshold(datExpr, powerVector = powers, dataIsExpr = T, networkType = "signed", verbose = 2, corFnc="bicor")

sizeGrWindow(9, 5)
#pdf(file='~/Documents/MSC_project/data/WGCNA/rld_soft_thresholding(bicor).pdf')
par(mfrow = c(1,2));
cex1 = 0.5;
 #Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
labels=powers,cex=cex1,col="red");


# this line corresponds to using an R^2 cut-off of h
#selected 0.9 as threshold to 
abline(h=0.90,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n", yaxt="n", ylim= c(0,7000),
main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
axis(2, at=seq(from = 0, to = 7000, by =1000),labels=seq(from = 0, to = 7000, by =1000), col.axis="black", las=2) 

dev.off()
```

#still many nodes with high connectivity.. p(k) proportion of nodes with connectivity k
```{r}
k=softConnectivity(datE=datExpr, corFnc = 'bicor', power=14, type = 'signed') 
#settings used to construct the network

pdf(file='~/Documents/MSC_project/data/WGCNA/check_scale_free_topo.pdf')
#sizeGrWindow(10,5)
par(mfrow=c(1,2))
hist(k, breaks=10, col = 'grey78')
scaleFreePlot(k, main="Check scale free topology\n")

dev.off()

```

```{r}
#will try step-by-step first, then can attempt blockwise modules function
softpower = 14
#plot adj in next section
adj= adjacency(datExpr,type = "signed", power = softpower, corFnc = 'bicor')

#topological overlap matrix: transfrom adj into TOM to minimise effects of noise and spurious associations
#The TOM describes how well connected the genes are in respect of how many neighbors they share. 
#idea of TOM is to count the direct connection strengths as well as connection strengths “mediated” by shared neighbors
#signed nowick: preserves the sign of the relationship between the nodes while taking account the connections through shared neighbors in the same way the standard unsigned TOM does.. unsure if this is used much..

TOM = TOMsimilarity(adj)

#save bicor TOM 
write.table(TOM, file = '/Users/martingordon/Documents/MSC_project/data/WGCNA/TOM_bicor.txt', sep ='\t', col.names = T, row.names = T)

#take the distance of the TOM matrix
dissTOM = 1-TOM

```

Plot the adjacency matrix
```{r}
gene_ids <-rownames(adj)

#plot adjacency matrix
adj <- matrix(adj, nrow=nrow(adj))
rownames(adj) <- gene_ids
colnames(adj) <- gene_ids

#topological overlap matrix 
#minimise noise and spurious associations, trandorm into TOM ans calculate corresponding similarity

#take random sample of 500 and plot 
heatmap_indices <- sample(nrow(adj), 500)

#we can still see here a subset of genes highly correlated... not ideal as ideally want sparseness but can see some genes v similiar correlation of expression
#command used to plot multiple adj.
#pdf(file='~/Documents/MSC_project/data/WGCNA/rld_bicor_adj_matrix_20thresh.pdf')
heatmap.2(t(adj[heatmap_indices, heatmap_indices]),
            col=redgreen(75),
            labRow=NA, labCol=NA, 
            trace='none', dendrogram='row',
            xlab='Gene', ylab='Gene',
            main='     Adjacency matrix (rld, bicor, power:20)',
            density.info='none', revC=TRUE)

#dev.off()
```

Module Definition; hierarchical clustering and tree cutting to define modules of co-expressed genes

```{r}
# Call the hierarchical clustering function from flashclust as faster than base hclust

geneTree = flashClust::hclust(as.dist(dissTOM), method = "average"); #avergae standard param; average dissimilarities between all objects
# Plot the resulting clustering tree (dendrogram)
#each verticle line corresponds to a gene, clusters corrwspond to 'leafs ' of genes
#dense branches indicate highly inter-connected, co-expressed genes
#module identification = identification of individual branches

#selecting the cut height leads to v different results, NB step
#use the dynamic tree cut package: bottom-up approach, merge branches evaluate (no of objects, gap between branches etc.),if meet criteria for being a modulle called seperate modules, otherwise merge
#can merge branches based on similiar eigengene values

sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity",
labels = FALSE, hang = 0.04)



# Module identification using dynamic tree cut: step 1: detection of clusters

#deepsplit; how finely branches should be split; higher value lmore smaller modules, lower value frwer larger modules
#PAM stage, turn PAM on or off; basically allows assigning more outlying objects to clusters
#This Partitioning Around Medoids-like step is based on average dissimilarities between unlabeled objects and the basic clusters detected in step 1 
#w/o PAM more grey (unassigned), with PAM dendogram can be difficult to interpret
#issue with tree cutting is its heuristic (potentially not optimal) for deciding if branch is a cluster


# want large modules, so we set the minimum module size relatively high:
minModuleSize = 30

dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM,
deepSplit = 2, pamStage = F, pamRespectsDendro = FALSE,
minClusterSize = minModuleSize);
table(dynamicMods) #29 modules detected

```

Plot dendogram, modules and traits
```{r}
# Convert numeric lables into colors
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)


# Convert traits to a color representation: white means low, red means high, grey means missing entry
traitColors = data.frame(numbers2colors(datTraits, signed = FALSE))
dim(traitColors)

# Plot the dendrogram and colors underneath
#sizeGrWindow(8,6)
#pdf(file = '/Users/martingordon/Documents/MSC_project/data/WGCNA/rld_gene_dend(bicor_20, deepsp = 1, PAM = T).pdf', wi = 9, he = 6)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05,
main = "Gene dendrogram and module colors") 


                    
table(rownames(datTraits) == rownames(datExpr))                  
                    
#dev.off()

```


Merge modules with similiar expression profiles since genes could be highly co-expressed
To calculate co-expression similarity of modules, calculate their eigengene value (a i-D vector that summarises the expression data of the genes that form a module, calculated from the first prinical component od the expression data)
Eigengenes represent module in mathematical calculations; can check for correlation with one another, can cluster, correlate with traits (i.e. case vs control)

```{r}
# Calculate eigengenes, match colours to dendogram
MEList = moduleEigengenes(datExpr, colors = dynamicColors)
MEs = MEList$eigengenes

# Calculate dissimilarity of module eigengenes
MEDiss = 1-cor(MEs);

# Cluster module eigengenes
METree = hclust(as.dist(MEDiss), method = "average");
# Plot the result

MEDissThres = 0.25 #correlation of 75% to merge. 

sizeGrWindow(7, 6)
pdf(file = '/Users/martingordon/Documents/MSC_project/data/new_WGCNA/module_eigengene_clustering.pdf', wi = 9, he = 6)
plot(METree, main = "Clustering of module eigengenes",
xlab = "", sub = "")
abline(h=MEDissThres, col = "red")

dev.off()

# Call an automatic merging function; merge modules with eigengenes similarity >=75%
merge = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 3)

# The merged module colors
mergedColors = merge$colors;
# Eigengenes of the new merged modules:
mergedMEs = merge$newMEs;

mergedMEs
#plot shows the orignal module definition by dynamic tree cut and then the merging of modules via correlation in eigengene values
sizeGrWindow(12, 9)
#
pdf(file = '/Users/martingordon/Documents/MSC_project/data/new_WGCNA/Cluster_dend(bicor_20,eign_thresh=.25).pdf', wi = 9, he = 6)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
c("Dynamic Tree Cut", "Merged dynamic"),
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05)

dev.off()

```

#plot heatmap of genes to visualise 
#light colours, little overlap, dark colours high overlap (= modules)
# v slow to plot..take subset of genes to speed up (take 4k)
#TOM in squares, rows are genes and columns and sorted by the clustering algorithm

```{r}
#heatmap colours
myheatcol = colorpanel(250,'red',"orange",'lemonchiffon')

nSelect = 2000 #intesnive, so take a subet of the expression matrix

# For reproducibility, we set the random seed
#set.seed(10);
select = sample(ncol(datExpr), size = nSelect);
selectTOM = dissTOM[select, select];

# no simple way of restricting a clustering tree to a subset of genes, so we must re-cluster.
selectTree = hclust(as.dist(selectTOM), method = "average")
selectColors = dynamicColors[select];


sizeGrWindow(9,9)
# Taking the dissimilarity to a power, makes the plot more informative by effectively changing the color palette; 
#setting the diagonal to NA also improves the clarity of the plot

plotDiss = selectTOM^9;
diag(plotDiss) = NA;

#takes long time to write out to pdf
pdf(file = '/Users/martingordon/Documents/MSC_project/data/new_WGCNA/gene_network_heatmap.pdf', wi = 9, he = 6)
TOMplot(plotDiss, selectTree, selectColors, main = "Network heatmap plot, selected genes", col= myheatcol)

dev.off()
```

Plot eigengene correlation networks: 
Look at relationships among the found modules. One can use the eigengenes as representative profiles and quantify module similarity by eigengene correlation. 

```{r}
plotEigengeneNetworks(mergedMEs, "", marDendro = c(0, 4, 1, 2), marHeatmap = c(3, 
    4, 1, 2), cex.lab = 0.8, xLabelsAngle = 90, col=)

# Plot the dendrogram
#pdf(file = '/Users/martingordon/Documents/MSC_project/data/new_WGCNA/eigengene_dendogram.pdf', wi = 9, he = 6)
plotEigengeneNetworks(mergedMEs, "Eigengene dendrogram", marDendro = c(0,4,2,0), marHeatmap = c(3,4,2,2),
plotHeatmaps = T)
#dev.off()
```

Use the merged module colors in mergedColors. Save the relevant variables for use in subsequent parts of the analysis

```{r}
# Rename to moduleColors
moduleColors = mergedColors
# Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50));
moduleLabels = match(moduleColors, colorOrder)-1;
MEs = mergedMEs
# Save module colors and labels for use in subsequent parts
save(MEs, moduleLabels, moduleColors, geneTree, file = "/Users/martingordon/Documents/MSC_project/data/new_WGCNA/network_construction.RData")
```

Module trait relationship plot 
Identify modules significantly associated with traits
#Correlate eigengene values with condition and assess gene significance (GS) measures; statistical significance between the oth node profile and the sample traits
#Each entry module eigengene correlated with the trait
```{r}
# Define numbers of genes and samples
nGenes = ncol(datExpr);
nSamples = nrow(datExpr);

head(datExpr)

#make into df for the module-trait analysis
#binary traits so encode as 1s and 0s
Control <- c(1,1,1,1, 0,0,0,0)
miR_Overexp <- c(0,0,0,0, 1,1,1,1)

datTraits <- data.frame(Control, miR_Overexp)


datTraits

# Recalculate MEs with color labels
#correlate with case/control status 
MEs0 = moduleEigengenes(datExpr, moduleColors)$eigengenes
MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, datTraits, use = "p"); #pearson correlation to assess correlation between merged eigengenes and traits
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);

MEs

# Form a data frame analogous to expression data that will hold the clinical traits.

samples = rownames(datExpr);
samples

dim(moduleTraitCor)


sizeGrWindow(10,6)
#pdf(file = '/Users/martingordon/Documents/MSC_project/data/new_WGCNA/MTR_plot.pdf')
# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
signif(moduleTraitPvalue, 1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCor)

par(mar = c(6, 8.5, 3, 3));
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor,
xLabels = names(datTraits),
yLabels = names(MEs),
ySymbols = names(MEs),
colorLabels = FALSE,
colors = blueWhiteRed(50),
textMatrix = textMatrix,
setStdMargins = FALSE,
cex.text = 0.5,
zlim = c(-1,1),
main = paste("Module-trait relationships"))

#dev.off()
```

Intramodular analysis; identifying genes with high GS and MM
Gene relationship to trait and important modules: Gene Significance and Module Membership
Quantify associations of individual genes with trait of interest by defining Gene Significance GS as (the absolute value of) the correlation between the gene and the trait. 
For each module, also define a quantitative measure of module membership MM as the correlation of the module eigengene and the gene expression profile. This allows us to quantify the similarity of all genes on the array to every module.

```{r}
#focus on modules associated with miR overexpression

miR_17 <- as.data.frame(datTraits$miR_Overexp);
names(miR_17) = "miR-17-92"

# names (colors) of the modules
modNames = substring(names(MEs), 3)

# quantify associations of individual genes with our trait of interest by defining Gene Significance GS as the absolute value of the correlation between the gene and the trait
#For each module, we also define a quantitative measure of module membership MM as the correlation of the module eigengene and the gene expression profile. 

geneModuleMembership = as.data.frame(cor(datExpr, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep=""); #p value for module membership of each gene

geneTraitSignificance = as.data.frame(cor(datExpr, miR_17, use = "p")); #correlation of each gene  with trait

GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples)); #the resultant p value

names(geneTraitSignificance) = paste("GS.", names(miR_17), sep="");

names(GSPvalue) = paste("p.GS.", names(miR_17), sep="");

```

GS:Correlating gene significance with clinical traits

```{r}
#average gene significance as a measure of module significance 
miR_17 <- c(0,0,0,0,1,1,1,1)

miR_ctrl <- c(1, 1,1, 1, 0,0,0,0)

GS2=as.numeric(cor(miR_17,datExpr, use="p"))
GeneSignificance=abs(GS2)


# Next module significance is defined as average gene significance.
ModuleSignificance=tapply(GeneSignificance, moduleColors, mean, na.rm=T)

#pdf(file = '/Users/martingordon/Documents/MSC_project/data/new_WGCNA/gene_sig_barplot.pdf')
plotModuleSignificance(GeneSignificance, moduleColors) #could alos use this to define pathway membership etc.


ModuleSignificance
dev.off()

#dev.off()
```


```{r}

# calculate the module membership values (aka. module eigengene based connectivity kME)

datKME <- signedKME(datExpr, MEs )
datKME #ie how correlated, anticorrelated genes are with module eigengenes


pdf(file = '/Users/martingordon/Documents/MSC_project/data/new_WGCNA/GSvsMMplot.pdf')

colorOfColumn = substring(names(MEs), 3)
colorOfColumn
par(mfrow = c(2, 2))
selectModules = c("pink", 'lightgreen', 'turquoise', 'lightcyan')
#par(mfrow = c(2, length(selectModules)/2))
for (module in selectModules) {
    column = match(module, colorOfColumn)
    print(column)
    restModule = dynamicColors == module
    verboseScatterplot(datKME[restModule, column], GS3[restModule], xlab = paste("MM ", 
        module, "module"), ylab = "GS.miR_Overexp", main = paste("kME.", module, 
        "vs. GS"), col = module,  bty="n" , abline = T, pch = 16 , cex = 1, cex.axis = 1, cex.main = 1)
}

dev.off()

```


```{r}
#calculate intramodular connectivty
connect.measures <- intramodularConnectivity(adj, moduleColors) #module colours vector equilivant to dynamic colours 

#gene significance value for miR-17-92 overexp - take abs value of GS1
GeneSignificance <- abs(GS1)


head(connect.measures)

#relationship between gene significance ina intramodular connectivity

pdf(file = '/Users/martingordon/Documents/MSC_project/data/new_WGCNA/GSvsIntraModConnect.pdf')

colorlevels=unique(moduleColors)
rm_colorlevels = c( "grey", "black", "white", "darkred", "lightyellow")

colorlevels <- colorlevels[! colorlevels %in% rm_colorlevels]
colorlevels

#sizeGrWindow(9,6)
pdf(file = '/Users/martingordon/Documents/MSC_project/data/new_WGCNA/GSvsIntraModConnect_filtered.pdf')
par(mfrow=c(2,as.integer(0.5+length(colorlevels)/2)))
par(mar = c(4,5,3,1))
for (i in c(1:length(colorlevels)))
{
whichmodule=colorlevels[[i]];
restrict1 = (moduleColors==whichmodule); #could I use this to restrict bootstrap analysis to corresponding module?
verboseScatterplot(connect.measures$kWithin[restrict1],
GeneSignificance[restrict1], col=moduleColors[restrict1],
main=whichmodule,
xlab = "Connectivity", ylab = "Gene Significance", abline = TRUE, cex.lab = 1, cex = 1, pch = 16, cex.main = 1)
}

#can see that the turquoise, pink, lightgreen and lightcyan hub genes (high connectivity) tend to have high gene significance (correlation with the trait)

#this shows how well connected each gene is within the module
#want genes highly connected with high trait significance

dev.off()
```

Write out results for genes with their module membership values

```{r}
names(datExpr)

names(datExpr)[moduleColors=="lightcyan"] 
names(datExpr)[moduleColors=="darkgreen"] 
names(datExpr)[moduleColors=="turquoise"] 
names(datExpr)[moduleColors=="pink"] 

#create geneinfo dataframe

geneInfo0 <- data.frame(Entrez_ID = names(datExpr), 
                       module_Colour = moduleColors,
                       geneTraitSignificance,
                       GSPvalue)

#order genes in df by significance for miR-17-92 overexpression
modOrder = order(-abs(cor(MEs, miR_Overexp, use = "p")));

for (mod in 1:ncol(geneModuleMembership))
{
oldNames = names(geneInfo0)
geneInfo0 = data.frame(geneInfo0, geneModuleMembership[, modOrder[mod]],
MMPvalue[, modOrder[mod]]);
names(geneInfo0) = c(oldNames, paste("MM.", modNames[modOrder[mod]], sep=""),
paste("p.MM.", modNames[modOrder[mod]], sep=""))
}

# Order the genes in the geneInfo variable first by module color, then by geneTraitSignificance
geneOrder = order(geneInfo0$module_Colour, -abs(geneInfo0$GS.miR.17.92));
geneInfo = geneInfo0[geneOrder, ]


#write df to file:
write.csv(geneInfo, file = "/Users/martingordon/Documents/MSC_project/data/new_WGCNA/geneInfo.csv", row.names = F)

```