---
title: "STRING_db_analysis"
author: "Martin Gordon"
date: "6/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## STRINGdb analysis of module genes

take the genes in the modulees/DEG interesction
pink upreg genes, turquoise downreg

```{r}
library(STRINGdb)
library(dplyr)
library(tibble)
library(networkD3)
library(visNetwork)
library(igraph)
library(knitr)
```


# Construct String_db network

```{r}
#read in list of differentially expressed genes 
#taken from DESeq2 analysis
up_df
string_up_df <- up_df[, c(2,5)]
string_up_df

down_df
string_down_df <- up_df[, c(2,5)]
string_down_df

#now subset these by the module overlap from cluster enrichment analysis
pinkmod_genes_filtered <- cyt_pinkf[["nodeData"]][["nodeName"]]
pinkmod_genes_filtered


string_up_df <- string_up_df[pinkmod_genes_filtered,]
#dplyr to convert rownmaes to col

string_up_df <- string_up_df %>% rownames_to_column('entrezgene_id')
string_up_df
```

#Now try run string on this set
#Map gene names to the string database using the 'map' method

```{r}
#version 11, h.sapiens, score threshold is 400 default
string_db <- STRINGdb$new( version="10", species=9606, score_threshold=400, input_directory="")

#couldnt map 8% of genes to identifiers.. Not found in db so remove
pink_mapped <- string_db$map(string_up_df, 'entrezgene_id', removeUnmappedRows = T )
pink_mapped

#only 119 hits (after removing those w/o equilivant STRING identifier) so plot the whole network

#pdf(file = "/Users/martingordon/Documents/MSC_project/data/STRING_module_analysis/pink_wholenetwork")
string_db$plot_network(pink_mapped$STRING_id)

#dev.off()
```
#Payload meechanism; 
#colour nodes in network by LFC change (red for overexpressed)
```{r}
# filter by p-value and add a color column
#i.e. green down-regulated gened and red for up-regulated genes)
pink_mapped_pval05 <- string_db$add_diff_exp_color( subset(pink_mapped, padj<0.05), logFcColStr="log2FoldChange" )

payload_id <- string_db$post_payload( pink_mapped_pval05$STRING_id, colors=pink_mapped_pval05$color )

png(file = "/Users/martingordon/Documents/MSC_project/data/STRING_module_analysis/pink_wholenetwork.png")
string_db$plot_network(pink_mapped$STRING_id, payload_id = payload_id)
dev.off()

```
Clustering igraph clustering algorithms to identify clusters 
```{r}
# get clusters
clustersList <- string_db$get_clusters(pink_mapped$STRING_id)


png(file = "/Users/martingordon/Documents/MSC_project/data/STRING_module_analysis/pink_networkclusters.png")

par(mfrow=c(2,2))
for(i in seq(1:4)){
string_db$plot_network(clustersList[[i]])
}

dev.off()
```
```{r}
#Get a table of the proteins in the db
string_proteins <- string_db$get_proteins() #protein names for the string ids
string_proteins


p <- dplyr::select(string_proteins, preferred_name) #extract protein name

proteins <- string_db$map(p, "preferred_name")
proteins


interactions <- string_db$get_interactions(pink_mapped$STRING_id) # protein interactions for genes within pink module

#map gene names to the string id's

for(i in 1:nrow(interactions)){
  row <- interactions[i,]
  from <- row$from
  to <- row$to
  from_ <- proteins[which(proteins$STRING_id == paste(from)),]
  to_ <- proteins[which(proteins$STRING_id == paste(to)),]
  from__ <- from_$preferred_name
  to__ <- to_$preferred_name
  interactions$from[i] <- from__
  interactions$to[i] <- to__
}

interactions

interactions <- interactions[,c(1,2,16)] #extract to,from and string score
interactions <- interactions[order(interactions$from),] #order
colnames(interactions) <- c("from", "to", "width")
interactions$width <- interactions$width/100 #line width for graph
```
#Creating plot

```{r}

pink_nodes <- data.frame(matrix(NA_character_, nrow = nrow(pink_mapped), ncol = 2))
pink_nodes[,1] <- pink_mapped$STRING_id
pink_nodes[,2] <- pink_mapped$STRING_id



colnames(pink_nodes) <- c("id", "label")

#map entrez_id to gene identifiers
for(i in 1:nrow(pink_nodes)){
  row <- pink_nodes[i,]
  id <- row$id
  label <- row$label
  id_ <- proteins[which(proteins$STRING_id == paste(id)),] 
  label_ <- proteins[which(proteins$STRING_id == paste(label)),]
  id__ <- id_$preferred_name
  label__ <- label_$preferred_name
  pink_nodes$id[i] <- id__
  pink_nodes$label[i] <- label__
}

#Create group column
pink_nodes[,"group"] <- "NA"

pink_nodes



hits <- pink_mapped$STRING_id
background <- hits 
string_db$set_background(background) #only the proteins from the module are included in the graph


#####   enrichment analysis

enrichment <- string_db$get_enrichment(pink_mapped$STRING_id, category = "KEGG")
head(enrichment, n=20)

annot <- string_db$get_annotations()
annot <- merge(annot, enrichment, by="term_id")
annot <- subset(annot, annot$term_id %in% enrichment$term_id)
annot <- subset(annot, annot$STRING_id %in% pink_mapped$STRING_id) #subset by genes in pink mod
annot <- subset(annot, annot$hits >= 4) #only keep terms with 4 or more hits in geneset



#map gene names to string id
for(i in 1:nrow(annot)){
  row <- annot[i,]
  id <- row$STRING_id
  id_ <- proteins[which(proteins$STRING_id == paste(id)),] #match string 
  id__ <- id_$preferred_name
  annot$STRING_id[i] <- id__
}


annot$STRING_id

outtable <- annot[,c(9,2,7,8)]
outtable
DT::datatable(outtable, rownames = FALSE)

#save enrichment results for genes in the module
write.table(outtable, "/Users/martingordon/Documents/MSC_project/data/STRING_module_analysis/pink_KEGGenrichment.txt", row.names = F, col.names = T, quote=F)


annot$Gene <- annot$STRING_id
#annot2 <- annot[,c(10,9)]



#map the KEGG pathway genes are associated to for the nodes
for(i in 1:nrow(pink_nodes)){
  row <- pink_nodes[i,]
  gene <- row$id
  df <- subset(annot, annot$STRING_id %in% paste(gene))
  vec <- as.vector(df$term_description)
  pink_nodes$group[i] <- paste(vec, collapse=",")
}


pink_nodes
annot
```


```{r}
#Vis Network - Interactive network for visualisation


pink_nodes <- na.omit(pink_nodes)
pink_nodes


#edges between
key <- c(interactions$from, interactions$to)
pink_nodes <- subset(pink_nodes, pink_nodes$id %in% key)



#Visualise as interactive network

visNetwork(pink_nodes, interactions, height = "1000px", width = "1000px", main="Pink Module PPI Network") %>% 
  visOptions(selectedBy = list(variable="group", multiple = TRUE), highlightNearest = list(degree = 2, enabled = T, hover = T, hideColor = "#F0FFFF"), 
   nodesIdSelection = list(enabled = T,
                           style = 'width: 200px; height: 26px;
                                 background: #f8f8f8;
                                 color: black;
                                 border:none;
                                 outline:none;')) %>% 
  addFontAwesome() %>%
  visPhysics(maxVelocity = 0.0001, 
             barnesHut = c(avoidOverlap=1),
             solver = "forceAtlas2Based",
               forceAtlas2Based = list(gravitationalConstant = -100)) %>%
  visLayout(improvedLayout = TRUE) %>%
visIgraphLayout(layout="layout_on_sphere") %>%
    visNodes(size = 50,shape = "dot", color = list(background = "#0085AF",border = "black",highlight = "#FF8000"), shadow = list(enabled = T, size = 10), font = c(size=30, bold = T)) %>%
  visEdges(smooth =F, 
    shadow = FALSE, width =1,
    color = list(color = "#B0C4DE", highlight = "#C62F4B"))

?visIgraphLayout
interactions$w/6

```


#do the same for turquoise module 

```{r}
down_df
string_down_df <- down_df[, c(2,5)]
string_down_df


#now subset these by the module overlap from cluster enrichment analysis
turq_genes_filtered <- cyt_turquoisef[["nodeData"]][["nodeName"]]
turq_genes_filtered


string_down_df <- string_down_df[turq_genes_filtered,]
#dplyr to convert rownmaes to col

string_down_df <- string_down_df %>% rownames_to_column('entrezgene_id')
string_down_df
```


```{r}

#couldnt map 7% of genes to identifiers.. Not found in db so remove
turq_mapped <- string_db$map(string_down_df, 'entrezgene_id', removeUnmappedRows = T )
turq_mapped

#only 174 hits (after removing those w/o equilivant STRING identifier) so plot the whole network

#pdf(file = "/Users/martingordon/Documents/MSC_project/data/STRING_module_analysis/pink_wholenetwork")
string_db$plot_network(turq_mapped$STRING_id)

#dev.off()
```

```{r}
# filter by p-value and add a color column
#i.e. green down-regulated gened and red for up-regulated genes)
turquoise_mapped_pval05 <- string_db$add_diff_exp_color( subset(turq_mapped, padj<0.05), logFcColStr="log2FoldChange" )

payload_id <- string_db$post_payload(turquoise_mapped_pval05$STRING_id, colors=turquoise_mapped_pval05$color )

png(file = "/Users/martingordon/Documents/MSC_project/data/STRING_module_analysis/turq_wholenetwork.png")
string_db$plot_network(turquoise_mapped_pval05$STRING_id, payload_id = payload_id)
dev.off()

```

#Clustering

```{r}
# get clusters
clustersList <- string_db$get_clusters(turq_mapped$STRING_id)

png(file = "/Users/martingordon/Documents/MSC_project/data/STRING_module_analysis/pink_networkclusters.png")

par(mfrow=c(2,2))
for(i in seq(1:4)){
string_db$plot_network(clustersList[[i]])
}

dev.off()
```


#Prepare interactive network 

```{r}
#Get a table of the proteins in the db
string_proteins <- string_db$get_proteins() #protein names for the string ids
string_proteins


p <- dplyr::select(string_proteins, preferred_name) #extract protein name

proteins <- string_db$map(p, "preferred_name")
proteins


interactions <- string_db$get_interactions(turq_mapped$STRING_id) # protein interactions for genes within turquoise module
#no observed interactions for these genes 

#map gene names to the string id's

for(i in 1:nrow(interactions)){
  row <- interactions[i,]
  from <- row$from
  to <- row$to
  from_ <- proteins[which(proteins$STRING_id == paste(from)),]
  to_ <- proteins[which(proteins$STRING_id == paste(to)),]
  from__ <- from_$preferred_name
  to__ <- to_$preferred_name
  interactions$from[i] <- from__
  interactions$to[i] <- to__
}

interactions

interactions <- interactions[,c(1,2,16)] #extract to,from and string score
interactions <- interactions[order(interactions$from),] #order
colnames(interactions) <- c("from", "to", "width")
interactions$width <- interactions$width/100 #line width for graph

interactions
```